A parametricity-based formalization of semi-simplicial and semi-cubical sets

### 2023-10-01 paper/paper.tex
Micro-changes

\begin{itemize}
  \item Sections~\ref{sec:ett} and~\ref{sec:itt} describe the construction at an informal level of discourse:
        \begin{itemize}
          \item In \ref{sec:ett}, we present it in informal extensional type theory where equational reasoning is left implicit.
          \item While reasoning in extensional type theory is similar to reasoning in set theory regarding how equality is implicitly handled, extensional type theory has two limitations. The first limitation is that it enforces the principle of Uniqueness of Identity Proofs and this is inconsistent with the Univalence principle, thus making it inexpressible in Homotopy Type Theory. The second limitation is that we want the construction to be formalizable in the Coq proof assistant whose underlying type theory is intensional. Section~\ref{sec:itt} thus rephrases the construction in (informal) intensional type theory.
        \end{itemize}
  \item Sections~\ref{sec:wf}, \ref{sec:le}, and \ref{sec:eqproperties} describe issues to be addressed in order to get a fully formal construction:
        \begin{itemize}
          \item The well-foundedness of the induction requires a special termination evidence which will be discussed in section~\ref{sec:wf}.
          \item The construction is indexed over integers and holds under some constraints on the range of these integers. There is a standard formalization dilemma in this kind of situation: either the constraints on the range are embedded in the construction so that the construction makes sense only on the corresponding range, or the construction is made first on a more general domain than needed but restricted to a smaller domain only at the time of use. We adopted the first approach, requiring however to let the construction be dependent on proofs of inequalities on natural numbers. We discuss how we deal with such dependencies in section~\ref{sec:le}.
          \item A number of obvious groupoid properties of equality as well as type isomorphisms have been left implicit in the informal definition. This is discussed in section~\ref{sec:eqproperties}.
        \end{itemize}
\end{itemize}

@@

The definition is dispatched over tables \ref{tab:coind}, \ref{tab:core}, \ref{tab:frames} and \ref{tab:faces}. The extra table~\ref{tab:coh} shows the proof of a coherence condition used in \ref{tab:faces}. Table \ref{tab:coind} describes a $\nu$-set in indexed form, as a coinductively-defined infinite sequence of types representing the limit of $n$-truncated $\nu$-sets. The definition of a $n$-truncated $\nu$-set is in turn described in \ref{tab:core}. In the infinite sequence of types representing a $\nu$-set, the $n$-th component is a type dependent over a $\fullframe$. The type $\fullframe$ represents a matching object. It is recursively defined in \ref{tab:frames}, using the auxiliary definitions of $\framep$, $\layer$ and $\painting$. The type $\layer$ relies on an operator of frame restriction $\restrf$ which is defined in \ref{tab:faces}. This restriction operator is defined using the auxiliary definitions of $\restrl$ and $\restrc$.

>>>>

\begin{itemize}
  \item Sections~\ref{sec:ett} and~\ref{sec:itt} describe the construction at an informal level of discourse:
        \begin{itemize}
          \item In \ref{sec:ett}, we present it in informal extensional type theory where equational reasoning is left implicit.
          \item While reasoning in extensional type theory is similar to reasoning in set theory regarding how equality is implicitly handled, extensional type theory has two limitations. The first limitation is that it enforces the principle of Uniqueness of Identity Proofs and this is inconsistent with the Univalence principle, thus making it inexpressible in Homotopy Type Theory. The second limitation is that we want the construction to be formalizable in the Coq proof assistant whose underlying type theory is intensional. Section~\ref{sec:itt} thus rephrases the construction in (informal) intensional type theory.
        \end{itemize}
  \item Sections~\ref{sec:wf}, \ref{sec:le}, and \ref{sec:eqproperties} describe issues to be addressed in order to get a fully formal construction:
        \begin{itemize}
          \item The well-foundedness of the induction requires a special termination evidence which will be discussed in section~\ref{sec:wf}.
          \item The construction is indexed over integers and holds under some constraints on the range of these integers. There is a standard formalization dilemma in this kind of situation: either the constraints on the range are embedded in the construction so that the construction makes sense only on the corresponding range, or the construction is made first on a more general domain than needed but restricted to a smaller domain only at the time of use. We adopted the first approach, requiring however the construction to be dependent on proofs of inequalities on natural numbers. We discuss how we deal with such dependencies in section~\ref{sec:le}.
          \item A number of standard groupoid properties of equality as well as type isomorphisms have been left implicit in the informal definition. This is discussed in section~\ref{sec:eqproperties}.
        \end{itemize}
\end{itemize}

@@

The definition is dispatched over tables \ref{tab:coind}, \ref{tab:core}, \ref{tab:frames} and \ref{tab:faces}. The extra table~\ref{tab:coh} shows the proof of a coherence condition used implicitly in \ref{tab:faces}. Table \ref{tab:coind} describes a $\nu$-set in indexed form, as a coinductively-defined infinite sequence of types representing the limit of $n$-truncated $\nu$-sets. The definition of a $n$-truncated $\nu$-set is in turn described in \ref{tab:core}. In the infinite sequence of types representing a $\nu$-set, the $n$-th component is a type dependent over a $\fullframe$. The type $\fullframe$ represents a matching object. It is recursively defined in \ref{tab:frames}, using the auxiliary definitions of $\framep$, $\layer$ and $\painting$. The type $\layer$ relies on an operator of frame restriction $\restrf$ which is defined in \ref{tab:faces}. This restriction operator is defined using the auxiliary definitions of $\restrl$ and $\restrc$.
